---
title: "SMURP Chamber expts"
format: 
  html:
    embed-resources: true
---

```{r}
library(tidyverse)
library(dygraphs)
library(patchwork)
library(purrr)
source("R/gems_functions.R")
theme_set(theme_classic())
```

# Metadata

### Natural nitrogen abundance and saturation concentration

* Natural abundance not used. Should be used for background correction?
* Argon and Nitrogen saturation determined for mean coastal salinity
and room temperature. ESL ambient seawater temp changed from 22 to 19C
over the three replicates.

Sensitivity of n2 and Ar sat to temp:

```{r}
sat_sens <- tibble(temp = c(23, 22, 19), sal = 31)
sat_sens <- sat_sens |> 
  mutate(n2_sat = cal_n2_sat(sal, temp),
         ar_sat = cal_ar_sat(sal, temp))
sat_sens
```

### Calculate constants

```{r}
lab_temp = 23
lab_sal = 31
frac_abd_15 <- 0.00367
frac_abd_14 <- 1 - frac_abd_15
nit_sat_umol <- cal_n2_sat(sal = lab_sal, temp = lab_temp) # calculate for lab temp, sal, and pressure
ar_sat_umol <- cal_ar_sat(sal = lab_sal, temp = lab_temp)

# Chamber area
dia <- 0.09
area <- pi * (dia / 2)^2
```

### Files, times, treatments, backgrounds

Times and backgrounds determined by inspection of plots below.

```{r}
files <- list.files("data", 
                    pattern = "^GEMS_2024.*\\.txt$",
                    full.names = TRUE)

experiment <- str_extract(files, "[^_]+_[^_]+(?=\\.txt$)")

metadata <- tibble(datafile = files,
                        experiment = as.factor(experiment),
                        replicate = as.factor(rep(1:3, each = 4)),
                        grass = grepl("grass", experiment), 
                        open = grepl("open", experiment),
)
```

# Initial plots for metadata

Plot raw and normalized data to find 
t0, t_anox, t_peak,
bg_29, bg_30, bg_40_28

## Read all files and produce plots

```{r}
read_all_gems <- function(file) {
  df <- read_gems(file)
  df_wide <- rga_wider(df)
  df_ratios <- rga_ratios(df_wide)
}

alldata_wide <- metadata |> 
  mutate(data_wide = map(datafile, read_all_gems))
```

General plots for mass 29 and 30 background

```{r}
pmap(list(alldata_wide$data_wide, 
          paste(alldata_wide$experiment, 
                alldata_wide$replicate, sep = "-")), 
     plot_wide)

```

Record t0, bg_29, bg_30

```{r}
t0 <- as.POSIXct(c("2024-08-19 15:30:00",
                   "2024-08-20 15:47:00",
                   "2024-08-21 15:15:00",
                   "2024-08-22 13:14:00",
                   "2024-09-11 13:33:00",
                   "2024-09-12 16:30:00",
                   "2024-09-13 12:34:00",
                   "2024-09-16 14:04:00",
                   "2024-09-24 13:27:00",
                   "2024-09-25 11:24:00",
                   "2024-09-26 13:05:00",
                   "2024-09-30 13:07:00"))

bg_29 <- c(7.4E-12, 
           1E-11, 
           8.1E-12, 
           1E-11, 
           1.1E-11, 
           1.1E-11, 
           1.1E-11, 
           9E-12, 
           9E-12, 
           9.1E-12, 
           9.0E-12, 
           9.0E-12)

bg_30 <- c(2.1E-12, 
           5.3E-12, 
           2.3E-12, 
           5.2E-12, 
           5.6E-12, 
           5.6E-12, 
           5.6E-12, 
           5.0E-12, 
           4.8E-12, 
           4.7E-12, 
           4.7E-12, 
           4.7E-12)

```

Argon-Nitrogen ratio plot for Ar/N2 background

```{r}
plot_ar <- function(df, title) {
    df %>% 
    select(timestamp, mass_40_28) %>% 
    filter(mass_40_28 < 0.1) |> 
    dygraph(main = title) %>% 
    dyOptions(logscale = TRUE,
              digitsAfterDecimal = 4)
}

pmap(list(alldata_wide$data_wide, 
          paste(alldata_wide$experiment, 
                alldata_wide$replicate, sep = "-")), 
     plot_ar)
```

Record Ar background

```{r}
bg_40_28 <- c(0.036,
              0.041,
              0.032,
              0.039,
              0.038,
              0.038,
              0.038,
              0.039,
              0.040,
              0.040,
              0.040,
              0.039)
```

add times and backgrounds to data

```{r}
alldata_wide <- alldata_wide |> 
  mutate(t0 = t0,
         bg_29 = bg_29,
         bg_30 = bg_30,
         bg_40_28 = bg_40_28)
```

## Calculate molarities

```{r}
alldata_norm <- alldata_wide |> 
  mutate(data_norm = pmap(list(data_wide, bg_29, bg_30, bg_40_28,
                               nit_sat_umol, ar_sat_umol, t0), 
                          norm_rga))
```

## Oxygen minimums

```{r}
plot_o2 <- function(df, title) {
    df %>% 
    select(et, mass_32) %>% 
    dygraph(main = title) %>% 
    dyOptions(logscale = TRUE,
              digitsAfterDecimal = 0)
}

pmap(list(alldata_norm$data_norm, 
          paste(alldata_norm$experiment, 
                alldata_norm$replicate, sep = "-")), 
     plot_o2)
```

Record O2 min time

```{r}
t_anox = c(20000, 12000, 16000, 7000, 
           28000, 17000, 6000, 8000, 
           15000, 12000, 7000, 8000)
```

## max DN

```{r}
plot_30 <- function(df, title) {
    df %>% 
    select(et, umol_30) %>% 
    dygraph(main = title)
}

pmap(list(alldata_norm$data_norm, 
          paste(alldata_norm$experiment, 
                alldata_norm$replicate, sep = "-")), 
     plot_30)
```

Record peak DN times

```{r}
t_peak = c(72000, 41000, 44000, 23000, 
           92000, 65000, 47000, 41000, 
           48000, 57000, 32000, 37000)
```

add times to data

```{r}
alldata_norm <- alldata_norm |> 
  mutate(t_anox = t_anox,
         t_peak = t_peak)
```

## Calculate final rates

### Ar loss rates

```{r}
calc_ar_loss_rate <- function(df, t_o2_min) {
  ar <- df |> 
    select(et, umol_40) |>  
    filter(et > t_o2_min - 1000,
           et < t_o2_min + 1000)
  
  coef(lm(umol_40 ~ et, data = ar))[2] * 3600
}

alldata_ar <- alldata_norm |> 
  mutate(ar_loss_init = pmap_dbl(list(data_norm, 10000),
                             calc_ar_loss_rate),
         ar_loss_20 = pmap_dbl(list(data_norm, .2 * t_peak),
                             calc_ar_loss_rate),
         ar_loss_anox = pmap_dbl(list(data_norm, t_anox),
                             calc_ar_loss_rate),
         ar_loss_max = pmap_dbl(list(data_norm, t_peak - 1000),
                             calc_ar_loss_rate),)
```

### DN Rates

DN rate at O2 minimum and max mass 30

```{r}
alldata_dn <- alldata_ar |> 
  mutate(dn_init = pmap_dbl(list(data_norm, 10000),
                                calc_rate),
         dn_20 = pmap_dbl(list(data_norm, .2 * t_peak),
                             calc_rate),
         dn_anox = pmap_dbl(list(data_norm, t_anox),
                                calc_rate),
         dn_max = pmap_dbl(list(data_norm, t_peak - 1000),
                                   calc_rate))
```

Ar loss corrected rates

```{r}
alldata <- alldata_dn |> 
  mutate(dn_init_ar_cor = dn_init - ar_loss_init,
        dn_20_ar_cor = dn_20 - ar_loss_20,
        dn_anox_ar_cor = dn_anox - ar_loss_anox,
        dn_max_ar_cor = dn_max - ar_loss_max)
```

Scale rates to umol m-2 d-1

```{r}
alldata <- alldata |> 
  mutate(across(starts_with("dn_"), \(x) x / area))
```

# Summary table


```{r}
alldata |> 
  select(experiment, replicate,
         t_anox, t_peak,
         dn_init_ar_cor,
         dn_20_ar_cor,
         dn_anox_ar_cor) |> 
  mutate(t_anox = t_anox/3600,
         t_peak = t_peak/3600) |> 
knitr::kable(digits = c(1, 1, 1, 1, 0, 0, 0),
             col.names = c("Experiment", "Replicate",
                           "Time to Anoxia [h]", "Time to Peak [h]",
                           "DN Rate (initial) [umol/m2/h]", 
                           "DN Rate (20% max) [umol/m2/h]",
                           "DN Rate (O2 min) [umol/m2/h]"))
```

# Summary plots

```{r}
init <- ggplot(alldata, aes(experiment, dn_init_ar_cor)) +
  geom_boxplot() +
  labs(title = "Initial DN rate (Ar corrected)",
       y = "15N2 production [umol/m2/h]") +
  theme(axis.title.x = element_blank(),  # Remove x-axis title
    axis.text.x = element_blank())

tw <- ggplot(alldata, aes(experiment, dn_20_ar_cor)) +
  geom_boxplot() +
  labs(title = "DN rate at 20% max (Ar corrected)",
       y = "15N2 production [umol/m2/h]") +
  theme(axis.title.x = element_blank(),  # Remove x-axis title
    axis.text.x = element_blank())

ox <- ggplot(alldata, aes(experiment, dn_anox_ar_cor)) +
  geom_boxplot() +
  labs(title = "DN rate at O2 min (Ar corrected)",
       y = "15N2 production [umol/m2/h]") +
  theme(axis.title.x = element_blank(),  # Remove x-axis title
    axis.text.x = element_blank())

max <- ggplot(alldata, aes(experiment, dn_max_ar_cor)) +
  geom_boxplot() +
  labs(title = "Max DN rate (Ar corrected)",
       y = "15N2 production [umol/m2/h]")

init / tw / ox / max + plot_layout(axis_titles = "collect")
```

# summary ANOVA

```{r}
a <- aov(dn_anox_ar_cor ~ experiment, data = alldata)
summary(a)
```

```{r}
a <- aov(dn_anox_ar_cor  ~ grass + open, data = alldata)
summary(a)
```

Grass vs no grass

```{r}
a <- aov(dn_anox_ar_cor ~ grass, data = alldata)
summary(a)
```

between replicates

```{r}
a <- aov(dn_anox_ar_cor ~ replicate, data = alldata)
summary(a)
```

# Chamber experiment descriptions

Chamber is cut-off 1 l polyethelyne nalgene bottle.
Fine screen glued into shoulder of bottle.
Cap has ports for inlet, outlet, temp, O2.
Recirculated flow using impeller pump, 
flowing past MIMS membrane. Label injected through luer port on return line.

## 2024-08-19 chamber sed closed sided

Chamber label experiment in sed-only, closed sided crate.

* Installed chamber. Sed-surface 10cm, chamber-surface 5cm.
* Started pump, purge bubbles. Large initial bubble below screen.
* Inject 5ml of argon-saturated 20 mmol/L 15NO3- solution.
* Power outage

## 2024-08-20 chamber sed open sided

Chamber label experiment in sed-only, open sided crate.

* Installed chamber. Sed-surface 10cm, chamber-surface 5cm.
* Started pump, purge bubbles. Large initial bubble below screen.
* Inject 5ml of argon-saturated 20 mmol/L 15NO3- solution.

Power outage at 8:30 AM. Past peak. Secondary peak of mass 46. 


## 2024-08-21 chamber grass open sided

## 2024-08-22 chamber grass closed sided

# Replicates

## 2024-09-11 chamber sed closed sided

Chamber label experiment in sed-only, closed sided crate.

* Installed chamber. Sed-surface 10cm, chamber-surface 6cm.
* Started pump, purge bubbles.
* Inject 5ml of argon-saturated 20 mmol/L 15NO3- solution.

## 2024-09-12 chamber sed open sided

Chamber label experiment in sed-only, open sided crate.

* Installed chamber. Sed-surface 10cm, chamber-surface 6cm.
* Started pump, purge bubbles.
* Inject 5ml of argon-saturated 20 mmol/L 15NO3- solution.

## 2024-09-11 chamber grass closed sided

Chamber label experiment in seagrass, closed sided crate.

* Installed chamber. Sed-surface 10cm, chamber-surface 6cm.
* Started pump, purge bubbles.
* Inject 5ml of argon-saturated 20 mmol/L 15NO3- solution.

## 2024-09-16 chamber grass open sided

Chamber label experiment in sed-only, closed sided crate.

* Installed chamber. Sed-surface 10cm, chamber-surface 6cm.
* Started pump, purge bubbles.
* Inject 5ml of argon-saturated 20 mmol/L 15NO3- solution.

# 3rd replicate

## 2024-09-24 chamber sed closed

Chamber label experiment in sed-only, closed sided crate.

* Installed chamber. Sed-surface 10cm, chamber-surface 6cm.
* Started pump, purge bubbles.
* Inject 5ml of argon-saturated 20 mmol/L 15NO3- solution.

## 2024-09-25 chamber sed open

Chamber label experiment in sed-only, closed sided crate.

* Installed chamber. Sed-surface 10cm, chamber-surface 6cm.
* Started pump, purge bubbles.
* Inject 5ml of argon-saturated 20 mmol/L 15NO3- solution.

## 2024-09-26 chamber grass closed

Chamber label experiment in grass, closed sided crate.

* Installed chamber. Sed-surface 10cm, chamber-surface 6cm.
* Started pump, purge bubbles.
* Inject 5ml of argon-saturated 20 mmol/L 15NO3- solution.

## 2024-09-30 chamber grass open

Chamber label experiment in grass, open sided crate.

* Installed chamber. Sed-surface 10cm, chamber-surface 6cm.
* Started pump, purge bubbles.
* Inject 5ml of argon-saturated 20 mmol/L 15NO3- solution.